Language Skills in an Ethnically Segmented Labour market: Estonia 1989 -- 2012

### 2018-08-04 paper/ijm.tex
Minor mistakes corrected to be in line with the responses to reviewers

The central results for unemployment regressions are presented in
\cref{tab:unemployment_estimation_by_sex_and_ethnic_same_sample}\footnote{We
 restrict both models to have the same number of observations as
 requested by a reviewer. The unrestricted results are provided in the Appendix (\Cref{tab:unemployment_estimation_by_sex_and_ethnic}).}.
The
table clearly indicates that Estonian language skills are associated
with approximately 5 percentage points lower unemployment for both
Russian men and women. We also see that the effect is not 
influenced in any major way by the individual characteristics--for men it is
virtually unchanged while for women the estimate falls by 30\%, from
6.5 to 4.5 percentage points.

@@

\begin{table}[htbp]
	\begin{center}
		\caption{Estimation results for log wage}
		\label{tab:wage_estimation_by_sex_and_ethnic_same_sample} %SKB 04.07.2018 14:22:Not yet updated only prepared
		%Accounting for occupation in model reduced the Estonian coefficient for women drastically (from 0.12 down 0.06)!
		%For other groups it is less
		\begin{tabular}{l | D{.}{.}{3} @{\qquad} D{.}{.}{3} @{\qquad} D{.}{.}{3} @{\qquad} | @{\qquad}
				D{.}{.}{3} @{\qquad} D{.}{.}{3} @{\qquad} D{.}{.}{3}}
			\toprule
			&                  \multicolumn{3}{c}{Men}                  &               \multicolumn{3}{c}{Women}                \\
			Estonians     & \multicolumn{1}{c}{\hspace{-12pt}1}   & \multicolumn{1}{l}{\hspace{12pt}2}   & \multicolumn{1}{l}{\hspace{12pt}3} & \multicolumn{1}{c}{\hspace{-12pt}1}   & \multicolumn{1}{l}{\hspace{12pt}2}   & \multicolumn{1}{c}{3}   \\\midrule
			Russian      & 0.112^{***}        & 0.078^{***}        & 0.043^{***}            & 0.094^{***}        & 0.039^{***}        & 0.016^{*}        \\
			          & (0.015)          & (0.015)          & (0.014)              & (0.011)          & (0.011)          & (0.009)          \\
			English      & 0.343^{***}        & 0.158^{***}        & 0.108^{***}            & 0.311^{***}        & 0.130^{***}        & 0.075^{***}        \\
			          & (0.013)          & (0.015)          & (0.013)              & (0.010)          & (0.010)          & (0.009)          \\
			\# Observations       & \multicolumn{1}{l}{21,785} & \multicolumn{1}{l}{21,785} & \multicolumn{1}{l}{21,785}     & \multicolumn{1}{l}{26,673} & \multicolumn{1}{l}{26,644} & \multicolumn{1}{c}{26,449} \\
			$R^{2}$      & 0.730           & 0.760           & 0.801               & 0.769           & 0.810           & 0.830           \\ \midrule
			Russians:     & \\
			Estonian      & 0.020           & -0.013           & -0.009               & 0.167^{***}        & 0.114^{***}        & 0.066^{***}        \\
			          & (0.018)          & (0.019)          & (0.018)              & (0.014)          & (0.014)          & (0.013)          \\
			English      & 0.248^{***}        & 0.152^{***}        & 0.103^{***}            & 0.223^{***}        & 0.089^{***}        & 0.054^{***}        \\
			          & (0.026)          & (0.027)          & (0.024)              & (0.021)          & (0.022)          & (0.018)          \\
			\# Observations       & \multicolumn{1}{l}{7978}  & \multicolumn{1}{l}{7978}  & \multicolumn{1}{l}{7978}      & \multicolumn{1}{l}{9904}  & \multicolumn{1}{l}{9895}  & \multicolumn{1}{c}{9697}  \\
			$R^{2}$      & 0.731           & 0.746           & 0.794               & 0.794           & 0.809           & 0.854           \\ \hline
			year dummies    & \V             & \V             & \V                 & \V             & \V             & \V             \\
			indiv. charact.  &              & \V             & \V                 &              & \V             & \V             \\
			workplace charact. &              &              & \V                 &              &              & \V             \\ \bottomrule
		\end{tabular}
		\begin{flushleft}
			\caption*{\legend \\ Standard errors (clustered on individuals) in parenthesis; \\ Individual characteristics are \modelTwo.  Workplace characteristics are \modelThreeAdd.  \restrictions}
		\end{flushleft}
	\end{center}
\end{table}

@@

	\caption*{\legend \\ Standard errors (clustered on individuals) in parenthesis. \\ Individual characteristics are \modelTwo. 	}
	\label{tab:age_group_trend}
\end{table}

@@

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Fri Jul 6 15:59:47 2018
\begin{table}[ht]
 \centering
 \caption{Average wage and time-trend corrected residual across
  location and education groups for Russian men}
 \begin{tabular}{llrrrr}
  \toprule
  & & &\multicolumn{2}{c}{Log Wage}\\
  Location & Education & \# Observations & EST 0 & EST 1 & Premium\\
  \midrule
  Tallinn & $<$ HS & 399 & -0.20 & -0.01 & 0.19 \\ 
  Tallinn & HS & 2164 & 0.07 & 0.04 & -0.03 \\ 
  Tallinn & College Degree & 610 & 0.15 & 0.30 & 0.15 \\ 
  Other & $<$ HS & 501 & -0.12 & -0.12 & -0.01 \\ 
  Other & HS & 2658 & -0.10 & -0.07 & 0.03 \\ 
  Other & College Degree & 429 & 0.17 & 0.11 & -0.06 \\ 
  \bottomrule
 \end{tabular}
 \begin{flushleft}
  Notes: \emph{Log Wage} refers to regression residual where log
  wage is explained by year fixed effects. \emph{EST 0} and
  \emph{EST 1} are mean values for those who cannot, and who can,
  speak Estonian; $\text{Premium} = \text{EST 1} - \text{EST 0}$.
  Education level \emph{< HS} refers to less than high school
  degree, \emph{HS} is the high school degree as the highest
  completed education.
 \end{flushleft}
 \label{tab:loc-lang-edu}
\end{table}

@@

The table indicates that Russian women need Estonian 26 percentage
points more at work than Russian men. The table also suggests that
Russian women are 16 percentage points more likely to be working in
Estonian-speaking environment when controlling for other factors.
Both figures are statistically significant (at 5\% and 10 \% level
respectively), although the standard errors are large. We can also see that
college education is associated by much more Estonian need and usage,
but in regions where Russian speakers are concentrated, the capital
Tallinn and the North-East, Estonian is less needed. In contrast, no
gender difference is visible for English needs.

@@

We document that all analysed language skills are associated with
unemployment and wage, however in a different way. Our most
intriguing observation is that the largest group of
Russian men, high school graduates, earn virtually no income premium if fluent in Estonian,
the majority and the official language of the country, while ability
to speak the language is related to a substantially
lower unemployment rate.
Our analysis supports \citet{YaoandOurs2015}, and \citet{Lindemann2013} findings that 
the labour market is segmented between ethnic and gender lines.
However, we find that ethnic segregation is less prevalent in the male
labour market.
This suggests for the majority of Russian-speaking men, Estonian
language gives a distinct advantage by facilitating hiring by
Estonian-dominated employers. We also confirm
\citet{leppik+vihalemm2015JofBaltStud} conclusion that
these
jobs are associated with little career mobility. There is also some
anecdotal evidence suggesting that offices are
typically manned by ethnic Estonians while manufacturing units have a
large share of Russian workers.
One can speculate that this outcome is related to a substantial population share
with negative attitudes toward Russians \citep{korts2009JofBaltStud}.
On a more positive tone, we find
that this may not be the case for highly educated workers.

The ethnic segregation seems to be less of an issue for women.
Unfortunately, neither \citet{Toomet2011} nor
\citet{leppik+vihalemm2015JofBaltStud} analyse men and women
separately. We find some evidence that female labour market is less
segregated along ethnic lines, and women are more frequently employed in jobs in which they need Estonian
language more, such as working in a direct contact with customers.
We also find evidence that over time Estonian is becoming more
important in the female labour market. Unfortunately, the current data
does not let us assess whether is due to decreasing segregation, or
economy--wide rise of the importance of Estonian language. No such trend is visible for
men suggesting that 20 years ---almost a generation--- after
establishing a nation-state and market economy, the
labour market integration is still not complete.

>>>>

The central results for unemployment regressions are presented in
\cref{tab:unemployment_estimation_by_sex_and_ethnic_same_sample}\footnote{We
 restrict both models to have the same number of observations as
 requested by a reviewer. The unrestricted results are provided in the Appendix (\Cref{tab:unemployment_estimation_by_sex_and_ethnic}).}.
The
table clearly indicates that Estonian language skills are associated
with an approximately 5 percentage points lower unemployment for both
Russian men and women. We also see that the effect is not 
influenced in any major way by the individual characteristics--for men it is
virtually unchanged while for women the estimate falls by 30\%, from
6.5 to 4.5 percentage points.

@@

\begin{table}[htbp]
	\begin{center}
		\caption{Estimation results for log wage}
		\label{tab:wage_estimation_by_sex_and_ethnic_same_sample} %SKB 04.07.2018 14:22:Not yet updated only prepared
		%Accounting for occupation in model reduced the Estonian coefficient for women drastically (from 0.12 down 0.06)!
		%For other groups it is less
		\begin{tabular}{l | D{.}{.}{3} @{\qquad} D{.}{.}{3} @{\qquad} D{.}{.}{3} @{\qquad} | @{\qquad}
				D{.}{.}{3} @{\qquad} D{.}{.}{3} @{\qquad} D{.}{.}{3}}
			\toprule
			&                  \multicolumn{3}{c}{Men}                  &               \multicolumn{3}{c}{Women}                \\
			Estonians     & \multicolumn{1}{c}{\hspace{-12pt}1}   & \multicolumn{1}{l}{\hspace{12pt}2}   & \multicolumn{1}{l}{\hspace{12pt}3} & \multicolumn{1}{c}{\hspace{-12pt}1}   & \multicolumn{1}{l}{\hspace{12pt}2}   & \multicolumn{1}{c}{3}   \\\midrule
			Russian      & 0.112^{***}        & 0.078^{***}        & 0.043^{***}            & 0.094^{***}        & 0.039^{***}        & 0.016^{*}        \\
			          & (0.015)          & (0.015)          & (0.014)              & (0.011)          & (0.011)          & (0.009)          \\
			English      & 0.343^{***}        & 0.158^{***}        & 0.108^{***}            & 0.311^{***}        & 0.130^{***}        & 0.075^{***}        \\
			          & (0.013)          & (0.015)          & (0.013)              & (0.010)          & (0.010)          & (0.009)          \\
			\# Observations       & \multicolumn{1}{l}{21,785} & \multicolumn{1}{l}{21,785} & \multicolumn{1}{l}{21,785}     & \multicolumn{1}{l}{26,673} & \multicolumn{1}{l}{26,644} & \multicolumn{1}{c}{26,449} \\
			$R^{2}$      & 0.730           & 0.760           & 0.801               & 0.769           & 0.810           & 0.830           \\ \midrule
			Russians:     & \\
			Estonian      & 0.020           & -0.013           & -0.009               & 0.167^{***}        & 0.114^{***}        & 0.066^{***}        \\
			          & (0.018)          & (0.019)          & (0.018)              & (0.014)          & (0.014)          & (0.013)          \\
			English      & 0.248^{***}        & 0.152^{***}        & 0.103^{***}            & 0.223^{***}        & 0.089^{***}        & 0.054^{***}        \\
			          & (0.026)          & (0.027)          & (0.024)              & (0.021)          & (0.022)          & (0.018)          \\
			\# Observations       & \multicolumn{1}{l}{7,978}  & \multicolumn{1}{l}{7,978}  & \multicolumn{1}{l}{7,978}      & \multicolumn{1}{l}{9,697}  & \multicolumn{1}{l}{9,697}  & \multicolumn{1}{c}{9,697}  \\
			$R^{2}$      & 0.731           & 0.746           & 0.794               & 0.794           & 0.809           & 0.854           \\ \hline
			year dummies    & \V             & \V             & \V                 & \V             & \V             & \V             \\
			indiv. charact.  &              & \V             & \V                 &              & \V             & \V             \\
			workplace charact. &              &              & \V                 &              &              & \V             \\ \bottomrule
		\end{tabular}
		\begin{flushleft}
			\caption*{\legend \\ Standard errors (clustered on individuals) in parentheses; \\ Individual characteristics are \modelTwo.  Workplace characteristics are \modelThreeAdd.  \restrictions}
		\end{flushleft}
	\end{center}
\end{table}

@@

	\caption*{\legend \\ Standard errors (clustered on individuals) in parentheses. \\ Individual characteristics are \modelTwo. 	}
	\label{tab:age_group_trend}
\end{table}

@@

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Fri Jul 6 15:59:47 2018
\begin{table}[ht]
 \centering
 \caption{Average wage and time-trend corrected residual across
  location and education groups for Russian men}
 \begin{tabular}{llrrrr}
  \toprule
  & & &\multicolumn{2}{c}{Log Wage}\\
  Location & Education & \# Observations & EST 0 & EST 1 & Premium\\
  \midrule
  Tallinn & $<$ HS & 399 & -0.20 & -0.01 & 0.19 \\ 
  Tallinn & HS & 2,164 & 0.07 & 0.04 & -0.03 \\ 
  Tallinn & College Degree & 610 & 0.15 & 0.30 & 0.15 \\ 
  Other & $<$ HS & 501 & -0.12 & -0.12 & -0.01 \\ 
  Other & HS & 2,658 & -0.10 & -0.07 & 0.03 \\ 
  Other & College Degree & 429 & 0.17 & 0.11 & -0.06 \\ 
  \bottomrule
 \end{tabular}
 \begin{flushleft}
  Notes: \emph{Log Wage} refers to regression residual where log
  wage is explained by year fixed effects. \emph{EST 0} and
  \emph{EST 1} are mean values for those who cannot, and who can,
  speak Estonian; $\text{Premium} = \text{EST 1} - \text{EST 0}$.
  Education level \emph{< HS} refers to less than high school
  degree, \emph{HS} is the high school degree as the highest
  completed education.
 \end{flushleft}
 \label{tab:loc-lang-edu}
\end{table}

@@

The table indicates that Russian women need Estonian 26 percentage
points more at work than Russian men. The table also suggests that
Russian women are 16 percentage points more likely to be working in
Estonian-speaking environment when controlling for other factors.
Both figures are statistically significant (at 5\% and 10 \% level
respectively), although the standard errors are large. We can also see that
college education is associated with much more Estonian need and usage,
but in regions where large amounts of Russian speakers live, the capital
Tallinn and the North-East, Estonian is less needed. In contrast, no
gender difference is visible for English needs.

@@

We document that all analysed language skills are associated with
unemployment and wage, however in different ways. Our most
intriguing observation is that the largest group of
Russian men, high school graduates, earn virtually no income premium if fluent in Estonian,
the majority and the official language of the country, while ability
to speak the language is related to a substantially
lower unemployment rate.
Our analysis supports \citet{YaoandOurs2015}, and \citet{Lindemann2013} findings that 
the labour market is segmented between ethnic and gender lines.
However, we find that ethnic segregation is less prevalent in the male
labour market.
This suggests for the majority of Russian-speaking men, Estonian
language gives a distinct advantage by facilitating hiring by
Estonian-dominated employers. We also confirm
\citet{leppik+vihalemm2015JofBaltStud} conclusion that
these
jobs are associated with little career mobility. There is also some
anecdotal evidence suggesting that offices are
typically manned by ethnic Estonians while manufacturing units have a
large share of Russian workers.
One can speculate that this outcome is related to a substantial population share
with negative attitudes toward Russians \citep{korts2009JofBaltStud}.
On a more positive tone, we find
that this may not be the case for highly educated workers.

The ethnic segregation seems to be less of an issue for women.
Unfortunately, neither \citet{Toomet2011} nor
\citet{leppik+vihalemm2015JofBaltStud} analyse men and women
separately. We find some evidence that female labour market is less
segregated along ethnic lines, and women are more frequently employed in jobs in which they need Estonian
language more, such as working in a direct contact with customers.
We also find evidence that over time Estonian is becoming more
important in the female labour market. Unfortunately, the current data
does not let us assess whether this is due to decreasing segregation, or
economy--wide rise of the importance of Estonian language. No such trend is visible for
men suggesting that 20 years ---almost a generation--- after
establishing a nation-state and market economy, the
labour market integration is still not complete.
